---
title: "Prediction Assignment Writeup"
author: "Srustijit Sahoo"
date: "27 May 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).
The main goal of the project is to predict the manner in which 6 participants performed some exercise as described above. This is the "classe" variable in the training set. The machine learning algorithm described here is applied to the 20 test cases available in the test data and the predictions are submitted in appropriate format

## Data Loading and Analysis
The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

We have downloaded the files into the working directory.

```{r}
setwd("C:/Users/Kanha/Desktop/Coursera")
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")

library(knitr)
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
library(corrplot)
set.seed(301)

inTrain  <- createDataPartition(training$classe, p=0.7, list=FALSE)

TrainSet <- training[inTrain, ]

TestSet  <- training[-inTrain, ]
dim(TrainSet)
dim(TestSet)

NZV <- nearZeroVar(TrainSet)
TrainSet <- TrainSet[, -NZV]
TestSet  <- TestSet[, -NZV]
dim(TrainSet)
dim(TestSet)

AllNA    <- sapply(TrainSet, function(x) mean(is.na(x))) > 0.95
TrainSet <- TrainSet[, AllNA==FALSE]
TestSet  <- TestSet[, AllNA==FALSE]
dim(TrainSet)
dim(TestSet)

TrainSet <- TrainSet[, -(1:5)]
TestSet  <- TestSet[, -(1:5)]
dim(TrainSet)
dim(TestSet)
```
After cleaning, we can see that the number of vairables for the analysis are now only 54.


## Prediction Model Building
Two well known methods will be applied to model the regressions (in the Train dataset) and the best one (with higher accuracy when applied to the Test dataset) will be used for the quiz predictions. The methods are: Random Forests and Decision Tree, as described below. A Confusion Matrix is plotted at the end of each analysis to better visualize the accuracy of the models.

#Random Forests
```{r}
set.seed(301)
controlRF <- trainControl(method="cv", number=3, verboseIter=FALSE)
modFitRandForest <- train(classe ~ ., data=TrainSet, method="rf",
                          trControl=controlRF)
modFitRandForest$finalModel

predictRandForest <- predict(modFitRandForest, newdata=TestSet)
confMatRandForest <- confusionMatrix(predictRandForest, TestSet$classe)
confMatRandForest

plot(confMatRandForest$table, col = confMatRandForest$byClass, 
     main = paste("Random Forest - Accuracy =",
                  round(confMatRandForest$overall['Accuracy'], 4)))
```

#Decision Tree
```{r}
set.seed(301)
modFitDecTree <- rpart(classe ~ ., data=TrainSet, method="class")
fancyRpartPlot(modFitDecTree)

predictDecTree <- predict(modFitDecTree, newdata=TestSet, type="class")
confMatDecTree <- confusionMatrix(predictDecTree, TestSet$classe)
confMatDecTree

plot(confMatDecTree$table, col = confMatDecTree$byClass, 
     main = paste("Decision Tree - Accuracy =",
                  round(confMatDecTree$overall['Accuracy'], 4)))
```


##Applying the selected Model to the Test Data
The accuracy of the two regression modeling methods above are:

Random Forest : 0.9968 and Decision Tree : 0.8291. In that case, the Random Forest model will be applied to predict the 20 quiz results (testing dataset) as shown below.
```{r}
predictTEST <- predict(modFitRandForest, newdata=testing)
predictTEST
```
